One-arm spirals in locally isothermal, radially structured self-gravitating discs

### 2014-11-05 paper/model.tex
backup

\subsection{Disc model and initial conditions}
%We consider a smooth disc which contains a surface density bump
%between $R\in[R_{1},R_{2}]$, mimicking mass built-up in a dead
%zone. 
We adopt a modified power-law disc with surface
density profile 
\begin{align}
  \Sigma(R) = \Sigma_\mathrm{ref} \left(\frac{R}{R_0}\right)^{-s}\times B(R;
  R_{1}, R_{2}, \epsilon, \delta R_d), 
\end{align}
where $R_0$ is a reference cylindrical radius and $s$ is the power-law
index describing the smooth disc, and $\Sigma_\mathrm{ref}$ is a
surface density scale chosen by specifying $Q_\mathrm{out}$,
the Keplerian Toomre parameter at $R=R_{2}$,
\begin{align}
  Q_\mathrm{out} = \left.\frac{c_s\Omega_k}{\pi G
    \Sigma}\right|_{R=R_{2}}. 
\end{align}

The bump function
$B(R)$ represents a surface density boost between
$R\in[R_{1},R_{2}]$ by a factor $\epsilon^{-1}>1$,
and $\delta R_d$ is the transition width between the bump and the
smooth disc. We choose 
\begin{align}
  &B(R) = f_1(R)\times f_2(R),\\
  &f_1(R) = \frac{1}{2}\left(1 - \epsilon\right)\left[1 +
    \tanh\left(\frac{R-R_{1}}{\Delta_1}\right)\right]  + \epsilon,\\
  &f_2(R) = \frac{1}{2}\left(1 - \epsilon\right)\left[1 -
    \tanh\left(\frac{R-R_{2}}{\Delta_2}\right)\right]  + \epsilon,
\end{align}
where $\Delta_1 = \delta R_d \times H(R_{1})$ and similarly for
$R_{2}$. The bump function mimicks mass built-up in a dead zone.

@@

Our fiducial parameter values are: $s=2$, $R_{1}=R_0$, $R_{2}=2R_0$,
$\epsilon=0.1$, $\delta R_d=5$, $h=0.05$ and
$Q_\mathrm{out}=2$. An example of the initial surface density and the
Toomre parameter (see\S\ref{wkb}) is shown in Fig. \ref{initial_surf}. 
Initially there is no vertical or radial velocity
($v_R = v_r = v_\theta = 0$). The azimuthal velocity is initialized to
satisfy centrifugal balance with pressure and gravity,
\begin{align}
  \frac{v_\phi^2}{r} = \frac{1}{\rho}\frac{\p p}{\p r} + \frac{\p
    \Phi_\mathrm{tot}}{\p r}
\end{align}
and similarly in 2D. The angular velocity is $\Omega = v_\phi/R$.

>>>>

\subsection{Disc model and initial conditions}
%We consider a smooth disc which contains a surface density bump
%between $R\in[R_{1},R_{2}]$, mimicking mass built-up in a dead
%zone. 
We adopt a modified power-law disc with surface
density profile 
\begin{align}
  \Sigma(R) = \Sigma_\mathrm{ref} \left(\frac{R}{R_0}\right)^{-s}\times B(R;
  R_{1}, R_{2}, \epsilon, \delta R), 
\end{align}
where $R_0$ is a reference cylindrical radius and $s$ is the power-law
index describing the smooth disc, and $\Sigma_\mathrm{ref}$ is a
surface density scale chosen by specifying $Q_\mathrm{out}$,
the Keplerian Toomre parameter at $R=R_{2}$,
\begin{align}
  Q_\mathrm{out} = \left.\frac{c_s\Omega_k}{\pi G
    \Sigma}\right|_{R=R_{2}}. 
\end{align}

The bump function
$B(R)$ represents a surface density boost between
$R\in[R_{1},R_{2}]$ by a factor $\epsilon^{-1}>1$,
and $\delta R$ is the transition width between the bump and the
smooth disc. We choose 
\begin{align}
  &B(R) = f_1(R)\times f_2(R),\\
  &f_1(R) = \frac{1}{2}\left(1 - \epsilon\right)\left[1 +
    \tanh\left(\frac{R-R_{1}}{\Delta_1}\right)\right]  + \epsilon,\\
  &f_2(R) = \frac{1}{2}\left(1 - \epsilon\right)\left[1 -
    \tanh\left(\frac{R-R_{2}}{\Delta_2}\right)\right]  + \epsilon,
\end{align}
where $\Delta_1 = \delta R \times H(R_{1})$ and similarly for
$R_{2}$. The bump function mimicks mass built-up in a dead zone.

@@

Our fiducial parameter values are: $s=2$, $R_{1}=R_0$, $R_{2}=2R_0$,
$\epsilon=0.1$, $\delta R=5$, $h=0.05$ and
$Q_\mathrm{out}=2$. An example of the initial surface density and the
Toomre parameter (see\S\ref{wkb}) is shown in Fig. \ref{initial_surf}. 
Initially there is no vertical or radial velocity
($v_R = v_r = v_\theta = 0$). The azimuthal velocity is initialized to
satisfy centrifugal balance with pressure and gravity,
\begin{align}
  \frac{v_\phi^2}{r} = \frac{1}{\rho}\frac{\p p}{\p r} + \frac{\p
    \Phi_\mathrm{tot}}{\p r}
\end{align}
and similarly in 2D. The angular velocity is $\Omega = v_\phi/R$.
